---
title: "Time Varying Effect"
description: |
author:
  - name: Jacob Simmering, PhD 
    url: https://jacobsimmering.com
    affiliation: University of Iowa
    affiliation_url: https://uiowa.edu
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkg-load}
library(tidyverse)
library(patchwork)
library(survival)
library(gtsummary)
```

## Data Load and Processing
Load the data from the network drive using the correct path for argon vs. local
work. 

```{r gen-outcomes}
if (Sys.info()["sysname"] == "Darwin") {
  root_dir <- "/Volumes/lss_jsimmeri_backup/data/tz-5ari-final"
} else {
  root_dir <- "/Shared/lss_jsimmeri_backup/data/tz-5ari-final"
}
model_data <- read_rds(glue::glue("{root_dir}/treated_model_data.rds"))
```

Reduce to outcomes:

```{r}
outcomes <- model_data %>%
  select(enrolid, drug, develops_pd, survival_time)
```

And load the matching data:

```{r load-matches}
tz_tam_matches <- read_rds(glue::glue("{root_dir}/matches/tz_tam.rds")) %>%
  mutate(enrolid = as.numeric(enrolid))
tz_5ari_matches <- read_rds(glue::glue("{root_dir}/matches/tz_5ari.rds")) %>%
  mutate(enrolid = as.numeric(enrolid))
tam_5ari_matches <- read_rds(glue::glue("{root_dir}/matches/tam_5ari.rds")) %>%
  mutate(enrolid = as.numeric(enrolid))
```

With the matches and `outcomes`, build the matched outcome files:

```{r make-paired-outcomes}
outcomes_tz_tam <- inner_join(outcomes, tz_tam_matches, by = c("enrolid"))
outcomes_tz_5ari <- inner_join(outcomes, tz_5ari_matches, by = c("enrolid"))
outcomes_tam_5ari <- inner_join(outcomes, tam_5ari_matches, by = c("enrolid"))
```

## Base Model and Schoenfeld Residuals

Fit the base model, same as fit in the main analysis, and use the Schoenfeld 
residuals to get an estimate of the time-varying hazard ratio:

```{r}
cph_tz_tam <- coxph(Surv(survival_time, develops_pd) ~ drug, 
      data = outcomes_tz_tam, 
      cluster = pair_id,
      robust = TRUE)
cph_tz_5ari <- coxph(Surv(survival_time, develops_pd) ~ drug, 
      data = outcomes_tz_5ari, 
      cluster = pair_id,
      robust = TRUE)
cph_tam_5ari <- coxph(Surv(survival_time, develops_pd) ~ drug, 
      data = outcomes_tam_5ari, 
      cluster = pair_id,
      robust = TRUE)
```

For TZ/DZ/AZ vs tamsulosin:

```{r}
tibble(
  x = cox.zph(cph_tz_tam)[1]$time,
  y = as.numeric(cox.zph(cph_tz_tam)[1]$y)
) %>%
  ggplot(aes(x = x / 365, y = exp(y))) + 
  geom_smooth(method = "loess") + 
  geom_hline(
    yintercept = 1
  ) + 
  geom_hline(
    yintercept = exp(coef(cph_tz_tam)[1]), linetype = 2
  ) + 
  geom_hline(
    yintercept = exp(confint(cph_tz_tam)[1]), linetype = 3
  ) +
  geom_hline(
    yintercept = exp(confint(cph_tz_tam)[2]), linetype = 3
  ) + 
  scale_y_log10() + 
  labs(x = "Years of Follow-Up", y = "Hazard Ratio") +
  theme_bw()
```

For TZ/DZ/AZ vs 5ARI:

```{r}
tibble(
  x = cox.zph(cph_tz_5ari)[1]$time,
  y = as.numeric(cox.zph(cph_tz_5ari)[1]$y)
) %>%
  ggplot(aes(x = x / 365, y = exp(y))) + 
  geom_smooth(method = "loess") + 
  geom_hline(
    yintercept = 1
  ) + 
  geom_hline(
    yintercept = exp(coef(cph_tz_5ari)[1]), linetype = 2
  ) + 
  geom_hline(
    yintercept = exp(confint(cph_tz_5ari)[1]), linetype = 3
  ) +
  geom_hline(
    yintercept = exp(confint(cph_tz_5ari)[2]), linetype = 3
  ) + 
  scale_y_log10() + 
  labs(x = "Years of Follow-Up", y = "Hazard Ratio") +
  theme_bw()
```

And tamsulosin vs 5ARI:

```{r}
tibble(
  x = cox.zph(cph_tam_5ari)[1]$time,
  y = as.numeric(cox.zph(cph_tam_5ari)[1]$y)
) %>%
  ggplot(aes(x = x / 365, y = exp(y))) + 
  geom_smooth(method = "loess") + 
  geom_hline(
    yintercept = 1
  ) + 
  geom_hline(
    yintercept = exp(coef(cph_tam_5ari)[1]), linetype = 2
  ) + 
  geom_hline(
    yintercept = exp(confint(cph_tam_5ari)[1]), linetype = 3
  ) +
  geom_hline(
    yintercept = exp(confint(cph_tam_5ari)[2]), linetype = 3
  ) + 
  scale_y_log10() + 
  labs(x = "Years of Follow-Up", y = "Hazard Ratio") +
  theme_bw()
```

## Step Function
Fit with a time-varying effect but use a step function for each year of 
follow-up (first year, second year, and so on to 10th or later year).

```{r}
cph_step_tz_tam <- survSplit(Surv(survival_time, develops_pd) ~ ., 
          data = outcomes_tz_tam,
          cut = seq(1, 10, 2) * 365,
          episode = "tgroup") %>%
  coxph(Surv(tstart, survival_time, develops_pd) ~ treatment:strata(tgroup),
        data = ., cluster = pair_id, robust = TRUE)

cph_step_tz_5ari <- survSplit(Surv(survival_time, develops_pd) ~ ., 
          data = outcomes_tz_5ari,
          cut = seq(1, 10, 2) * 365,
          episode = "tgroup") %>%
  coxph(Surv(tstart, survival_time, develops_pd) ~ treatment:strata(tgroup),
        data = ., cluster = pair_id, robust = TRUE)

cph_step_tam_5ari <- survSplit(Surv(survival_time, develops_pd) ~ ., 
          data = outcomes_tam_5ari,
          cut = seq(1, 10, 2) * 365,
          episode = "tgroup") %>%
  coxph(Surv(tstart, survival_time, develops_pd) ~ treatment:strata(tgroup),
        data = ., cluster = pair_id, robust = TRUE)
```

The estimated time-varying HR:

```{r}
tbl_regression(cph_step_tz_tam, exponentiate = TRUE)
tbl_regression(cph_step_tz_5ari, exponentiate = TRUE)
tbl_regression(cph_step_tam_5ari, exponentiate = TRUE)
```

The same data graphically but overlaid with the main model estimates:

```{r}
cph_step_tz_tam %>%
  broom::tidy() %>%
  mutate(
    period = stringr::str_extract(term, "=[0-9]*"),
    period = as.numeric(stringr::str_replace(period, "=", "")),
    period = period - 1
  ) %>%
  ggplot(aes(x = period, y = exp(estimate),
             ymin = exp(estimate - 1.96 * robust.se),
             ymax = exp(estimate + 1.96 * robust.se))) +
  geom_point() + 
  geom_linerange() +
  geom_hline(
    yintercept = exp(coef(cph_tz_tam)[1]), linetype = 2
  ) + 
  geom_hline(
    yintercept = exp(confint(cph_tz_tam)[1]), linetype = 3
  ) +
  geom_hline(
    yintercept = exp(confint(cph_tz_tam)[2]), linetype = 3
  ) + 
  scale_y_log10() +
  labs(x = "Years of Follow-Up", y = "Estimated HR by Year") +
  theme_bw()

cph_step_tz_5ari %>%
  broom::tidy() %>%
  mutate(
    period = stringr::str_extract(term, "=[0-9]*"),
    period = as.numeric(stringr::str_replace(period, "=", "")),
    period = period - 1
  ) %>%
  ggplot(aes(x = period, y = exp(estimate),
             ymin = exp(estimate - 1.96 * robust.se),
             ymax = exp(estimate + 1.96 * robust.se))) +
  geom_point() + 
  geom_linerange() +
  geom_hline(
    yintercept = exp(coef(cph_tz_5ari)[1]), linetype = 2
  ) + 
  geom_hline(
    yintercept = exp(confint(cph_tz_5ari)[1]), linetype = 3
  ) +
  geom_hline(
    yintercept = exp(confint(cph_tz_5ari)[2]), linetype = 3
  ) + 
  scale_y_log10() +
  labs(x = "Years of Follow-Up", y = "Estimated HR by Year") +
  theme_bw()

cph_step_tam_5ari %>%
  broom::tidy() %>%
  mutate(
    period = stringr::str_extract(term, "=[0-9]*"),
    period = as.numeric(stringr::str_replace(period, "=", "")),
    period = period - 1
  ) %>%
  ggplot(aes(x = period, y = exp(estimate),
             ymin = exp(estimate - 1.96 * robust.se),
             ymax = exp(estimate + 1.96 * robust.se))) +
  geom_point() + 
  geom_linerange() +
  geom_hline(
    yintercept = exp(coef(cph_tam_5ari)[1]), linetype = 2
  ) + 
  geom_hline(
    yintercept = exp(confint(cph_tam_5ari)[1]), linetype = 3
  ) +
  geom_hline(
    yintercept = exp(confint(cph_tam_5ari)[2]), linetype = 3
  ) + 
  scale_y_log10() +
  labs(x = "Years of Follow-Up", y = "Estimated HR by Year") +
  theme_bw()
```
